Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
Polish power cmp
Now the summary of the findings is better. I mention that conventional test is still the recommended choice if the strict correctness of the model assumption is of particular interest.

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

@@

For both the non-linearity pattern and the heteroskedasticity pattern, the power curves of appropriate conventional tests are constantly greater than the power curves of visual tests and other inappropriate conventional tests. At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting $H_0$, while only a few visual tests would result in a rejection. Similarly, at $log(E) = 3$ for heteroskedasticity model, the BP test has around 25% power but the non-constant variance visual pattern is almost unobserved from the residual plot and the power of the visual test is close to the significance level at this point.

Further, the power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us. Visual test have much higher tolerance of the residual departures than the conventional test. Since fail to reject $H_0$ in a visual test usually means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test. Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible. Besides, the sensitivity of the conventional test could also distract and discourage analysts from finding simple but good linear approximation to the data. The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful. This may limit the popularity of conventional tests in residual diagnostics among analysts.

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 0, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

>>>>

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.

@@

For non-linearity patterns, the power curve of RESET test climbs aggressively from
`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is much more sensitive to weak non-linear structure. Meanwhile, no noticeable residual departures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is constantly greater than the power of visual test. At $log_e(E) \approx 2.5$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$. Similarly, the visual feature is nearly unobservable from the example residual plots at this level of effect size.

The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features. This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "right-triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The visual pattern becomes clearer as $log_e(E)$ increases. At $log_e(E) = 5$, the power of visual tests for both patterns reaches almost 100%.

Power curves of inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed in this paper. Visual test have much higher tolerance of the residual departures than the conventional test. Since fail to reject $H_0$ in a visual test literally means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test. Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible. Besides, the sensitivity of the conventional test to weak residual departures could also distract and discourage analysts from finding simple but good linear approximation to the data. The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful to analysts and decision makers. In contrast, if the strict correctness of the model assumption is of particular interest, conducting a conventional test is still the recommended choice based on the findings.

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.tex
Polish power cmp
Now the summary of the findings is better. I mention that conventional test is still the recommended choice if the strict correctness of the model assumption is of particular interest.

We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of predictors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary predictors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.

@@

For both the non-linearity pattern and the heteroskedasticity pattern,
the power curves of appropriate conventional tests are constantly
greater than the power curves of visual tests and other inappropriate
conventional tests. At \(log(E) = 1\) for non-linearity model, where the
residual departures are unnoticeable from the corresponding residual
plot, the RESET test has above 25\% of chance of rejecting \(H_0\),
while only a few visual tests would result in a rejection. Similarly, at
\(log(E) = 3\) for heteroskedasticity model, the BP test has around 25\%
power but the non-constant variance visual pattern is almost unobserved
from the residual plot and the power of the visual test is close to the
significance level at this point.

Further, the power of all inappropriate tests show improvement as the
effect increases but at a lower rate than the visual test in both
scenarios. This coincides the point made by \citet{cook1982residuals}
that residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs
significantly from visual tests in two regression diagnostics scenarios
designed by us. Visual test have much higher tolerance of the residual
departures than the conventional test. Since fail to reject \(H_0\) in a
visual test usually means that there are no obvious visual discoveries
found in the residual plot, analysts and the general public as the
consumers of the output may not be convinced of the existence of
significant residual departures in spite of the rejection of \(H_0\)
given by the conventional test. Even if the rejection is accepted, the
model violation may be considered as impactless due to the fact that
they are not clearly visible. Besides, the sensitivity of the
conventional test could also distract and discourage analysts from
finding simple but good linear approximation to the data. The rejection
of \(H_0\) because of human acceptable and negligible residual
departures is not practically meaningful and useful. This may limit the
popularity of conventional tests in residual diagnostics among analysts.

>>>>

We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of predictors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary predictors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(x\) and \(x^2\) in the auxiliary regression.

@@

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 7\% to 85\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is much more sensitive to weak non-linear
structure. Meanwhile, no noticeable residual departures can be spotted
from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is constantly greater than the power of visual test. At
\(log_e(E) \approx 2.5\), where the power curve of the visual test
remains at a low level, the Breusch-Pagan test has around 50\% chance of
rejecting \(H_0\). Similarly, the visual feature is nearly unobservable
from the example residual plots at this level of effect size.

The power of visual test arises steadily as \(log_e(E)\) increases from
3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 3.5\), a
weak ``S-shape'' and a weak ``right-triangle'' shape are presented in
Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
The visual pattern becomes clearer as \(log_e(E)\) increases. At
\(log_e(E) = 5\), the power of visual tests for both patterns reaches
almost 100\%.

Power curves of inappropriate tests show improvement as the effect
increases but at a lower rate than the visual test in both scenarios.
This coincides the point made by \citet{cook1982residuals} that
residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs
significantly from visual tests in two regression diagnostics scenarios
designed in this paper. Visual test have much higher tolerance of the
residual departures than the conventional test. Since fail to reject
\(H_0\) in a visual test literally means that there are no obvious
visual discoveries found in the residual plot, analysts and the general
public as the consumers of the output may not be convinced of the
existence of significant residual departures in spite of the rejection
of \(H_0\) given by the conventional test. Even if the rejection is
accepted, the model violation may be considered as impactless due to the
fact that they are not clearly visible. Besides, the sensitivity of the
conventional test to weak residual departures could also distract and
discourage analysts from finding simple but good linear approximation to
the data. The rejection of \(H_0\) because of human acceptable and
negligible residual departures is not practically meaningful and useful
to analysts and decision makers. In contrast, if the strict correctness
of the model assumption is of particular interest, conducting a
conventional test is still the recommended choice based on the findings.
